\section{Series}

\subsection{Definition}
\begin{definition}
    Let $(V, \norm{\cdot})$ be a normed vector space and let $a: \N \to V$ be a sequence in $V$. Let $K \in \N$. We say that a series
    $$\sum_{n=K}^{\infty}a_n$$
    is \emph{convergent} if the associatedd \emph{sequence of partial sums} $S_k:\N \to V$, i.e. the sequece $(S_K^n)_{n\in\N}$ converges. The term $S_K^n$ is, for $n \in \N$, defined as
    $$S_K^n:= \sum_{k=K}^na_k$$

    If $K = 0$, we ussually jusst write $S^n$ or even $S_n$ instead of $S_0^n$.

    If the series $\sum_{n=K}^\infty a_n$ is convergent, the \emph{value} of the series is by defintion equal to the limit of  the sequence of partial sums, i.e.
    $$\sum_{k=K}^\infty a_k := \limn S_k^n = \limn \sum_{k=K}^\infty a_k$$
\end{definition}

\subsection{Geometric series}
\begin{proposition}
    Let $a \ne 1$ and $n \in \N$. Then
    $$\sum_{k=0}^n a^k = \frac{1-a^{n+1}}{1-a}.$$
\end{proposition}
\begin{proof}
    We consider
    \begin{align*}
        (1-a)\sum_{k=0}^n a^k &= \sum_{k=0}^n a^k - a\sum_{k=0}^n a^k  \\
                              &= \sum_{k=0}^n a^k - \sum_{k=0}^n a^{k+1} \\
                              &= \sum_{k=0}^n a^k - \sum_{k=1}^{n+1} a^k \\
                              &= 1 - a^{n+1}
    \end{align*}
\end{proof}

\begin{proposition}[Geometric series]
    Let $a \in (-1,1)$. Then the series
    $$\sum_{k=0}^\infty a^k$$
    is convergent and has the value
    $$\sum_{k=0}^\infty a^k = \frac{1}{1-a}.$$
\end{proposition}

\subsection{The harmonic series}
\begin{proposition}[Harmonic series]
    The series 
    $$\sum_{k=1}^\infty \frac{1}{k}$$
    diverges.
\end{proposition}

\subsection{The hyperharmonic series}
\begin{proposition}[Hyperharmonic series]
    Let $p > 1$. Then the series 
    $$\sum_{k=1}^\infty \frac{1}{k^p}$$
    converges.
\end{proposition}

\begin{example}
    Here is an example of a series taking values in the normed vector space $(\R^2, \norm{\cdot})$:
    $$\sum_{k=1}^\infty \left(\frac{1}{k^2}, \left(\frac{1}{2}\right)^k\right)$$
\end{example}

\subsection{Only the tail matters for convergence}
\begin{lemma}
    Let $(V, \norm{\cdot})$ be a normed vector space and let $a: \N \to V$ be a sequence taking values in $V$. Let $K,L \in \N$. The series
    $$\sum_{n=K}^\infty a_n$$
    is conovergent is and only if the series
    $$\sum_{n=L}^\infty a_n$$
    is convergent. Moreover, if either the series converges, and $K < L$, then
$$\sum_{n=K}^\infty a_n = \sum_{n=K}^{L-1} + \sum_{n=L}^\infty a_n.$$
\end{lemma}

\begin{proposition}
    Let $a: \N \to V$ be a sequence, let $M \in \N$ and assume that the series
    $$\sum_{k=M}^\infty a_k$$
    is convergent. Then
    $$\lim_{m\to\infty} \sum_{k=m}^\infty a_k = 0.$$
\end{proposition}

\begin{proposition}[Index shift for series]
    Let $a: \N \to V$ be a sequence, let $M \in \N$ and let $\ell \in \N$. Then the series
    $$\sum_{k=M}^\infty a_k$$
    converges if and only if the series
    $$\sum_{k=M}^\infty a_{k+\ell}$$
    converges. Moreoever, if either series converges, then
    $$\sum_{k=M}^\infty a_{k+\ell} = \sum_{k=M+\ell}^\infty a_{k}.$$
\end{proposition}

\subsection{Divergence test}
\begin{proposition}
    Let $(V,\norm{\cdot})$ be a normed vector space, and let $a:\N \to V$ be a sequence in $V$. Suppose the series $\displaystyle\sum_{n=0}^{\infty}a_n$ is convergent. Then
    $$\limn a_n = 0.$$
\end{proposition}
\begin{proof}[Proof]
    Suppose the series $\displaystyle\sum_{n=0}^{\infty}a_n$ is convergent to $L \in V$. Then
    $$a_n = S_n - S_{n-1}$$
    where $S_n$ denote the partial sum $\displaystyle\sum_{k=0 }^{n }a_k$. Because $S_n$ and $S_{n-1}$ are both convergent to $L$, the sequence $(a_n)$ is convergent as well and converges to $L - L = 0$.
\end{proof}

\begin{theorem}[Divergence test]
    Let $(V, \norm{\cdot})$ be a normed vector space and let $a: \N \to V$ be a sequence in $V$. Suppose the limit $\limn a_n$ does not exist or is not equal to $0$. Then the series
    $$\sum_{n=0}^\infty a_n$$
    is divergent.
\end{theorem}

\subsection{Limit laws for series}
\begin{theorem}[Limit laws for series]
    Let $(V, \norm{\cdot})$ be a normed vector space and let $a,b: \N \to V$ be sequences in $V$. Suppose the series
    $$\sum_{n=0}^\infty a_n \qquad \text{and} \qquad \sum_{n=0}^\infty b_n$$
    are convergent. Suppose $\lambda \in \R$. Then
    \begin{enumerate}
        \item The series
            $$\sum_{n=0}^{\infty}(a_n + b_n)$$
            is converget and converges to
            $$\sum_{n=0}^\infty a_n + \sum_{n=0}^\infty b_n.$$
        \item The series 
            $$\sum_{n=0}^\infty \lambda a_n$$
            is convergent and converges to
            $$\lambda \sum_{n=0}^\infty a_n.$$
    \end{enumerate}
\end{theorem}
